---
title: "rOmnidrive: User Guide"
author: "Pedro J. Aphalo"

date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This package provides a wrapper on the OmniDriver from Ocean Optics. The 
runtime can be downloaded and installed for free. Please, use the Java
documentation provided by Ocean Optics. The names of all functions are
preserved but change from camel case as in `getApiVersion()` to a style
easier to read and more usual in R. For this example `get_api_version()`.

## Using rOmniDriver

The first step is to load the package, which automatically loads the 
OminiDriver Java library if it is installed in the computer.
We need then to initialize the API by creating a wrapper object in Java.
Optionally we can query the API version. Updates to OmniDriver
report an earlier API version if the interface has not changed.

```{r}
library(rOmniDriver)
w <- init_api()
get_api_version(w)
```

The next step is to open connections to all Ocean Optics spectrometers
directly connected through USB to the local computer. This function does
not 'see' spectrometers connected through a LAN. It returns the number of
spectrometers found. The spectrometers are accessed through and index 
(integer number) in the range 0 to $n - 1$ where $n$ is the number
of connected spectrometers.

```{r}
n <- open_all_spectrometers(w)
n
```

In what follows we will explicitly access the first spectrometer, with
index equal 0. We use `0L` to indicate that we mean a _long integer_.

```{r, eval=n>0L}
get_name(jwrapper = w, sr.index = 0L)
get_serial_number(jwrapper = w, sr.index = 0L)
```

As the default for `sr.index` is `0L` and arguments can be recognized by
position, the code in the next chunk is equivalent to the previous one.

```{r, eval=n>0L}
get_name(w)
get_serial_number(w)
```

We can query for a description of the optical bench. In this case after
obtaining the object describing the `bench` we directly use the Java
functions defined in OmniDriver Java code. We use `$` to mean 'call the
Java method' of this name. As this are the Java methods, their names use
camel case.

```{r, eval=n>0L}
bench <- get_bench(jwrapper = w, sr.index = 0L)
bench$getGrating()
bench$getFilterWavelength()
bench$getSlitSize()
```

```{r, eval=n>0L}
n_ch <- get_number_of_channels(jwrapper = w, sr.index = 0L)
n_ch
```

We can also query about other characteristics of the spectrometer, including
the limits of the range of accepted integration times.

```{r, eval=n>0L}
get_maximum_integration_time(jwrapper = w, sr.index = 0L)
get_minimum_integration_time(jwrapper = w, sr.index = 0L)
get_maximum_intensity(jwrapper = w, sr.index = 0L)
wl <- get_wavelengths(jwrapper = w, sr.index = 0L, ch.index = 0L)
head(wl)
get_firmware_model(jwrapper = w, sr.index = 0L) # API >= 2.40 only
get_firmware_version(jwrapper = w, sr.index = 0L)
get_number_of_pixels(jwrapper = w, sr.index = 0L, ch.index = 0L)
get_number_of_dark_pixels(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

Some _features_ are available only in some models of spectrometers. Before
using these features it is necessary to check whether they are available or
not.

The first example is a temperature sensor on the controller board. In this
case there is only one method available, which returns the board temperature
in degrees centigrades.

```{r, eval=n>0L}
if (is_feature_supported_board_temperature(jwrapper = w, sr.index = 0L)) {
  controller.board <- 
    get_feature_controller_board_temperature(jwrapper = w, sr.index = 0L)
  controller.board$getBoardTemperatureCelsius()
} else {
  print("feature not supported")
}
```


The next example, is for a similar temperature sensor in the CCD detector
board. 

```{r, eval=n>0L}
if (is_feature_supported_detector_temperature(jwrapper = w, sr.index = 0L)) {
  detector.board <- 
    get_feature_detector_board_temperature(jwrapper = w, sr.index = 0L)
  detector.board$getDetectorTemperatureCelsius()
} else {
  print("feature not supported")
}
```

Another feature is switching between 'normal' and 'raw' spectral data. Again
we have a single method, but in this case accepting an argument.

```{r, eval=n>0L}
if (is_feature_supported_spectrum_type(jwrapper = w, sr.index = 0L)) {
  spectrum.type <- get_feature_spectrum_type(jwrapper = w, sr.index = 0L)
  spectrum.type$setSpectrumType(1L) # raw
  spectrum.type$setSpectrumType(0L) # normal
} else {
  print("feature not supported")
}
```

The integration is set in microseconds. At least in some spectrometer
types, out-of-range values are silently ignored. Use the functions
described above to find the valid range, or use `get_integration_time()`
to confirm that the set method has succeeded.

```{r, eval=n>0L}
set_integration_time(jwrapper = w, time.usec = 5e6, sr.index = 0L, ch.index = 0L)
get_integration_time(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

Boxcar smoothing is a running averaging along the pixels (wavelengths). It is
possible to set the number of pixels as a half width. (Using boxcar smoothing
reduces the noise reducing the effective wavelength resolution but maintains
the temporal resolution.)

```{r, eval=n>0L}
set_boxcar_width(w, half.width.px = 0L, sr.index = 0L, ch.index = 0L)
get_boxcar_width(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

Setting the number of scans to average to a number larger than one results
in the pixel-by-pixel averaging of several scans. (Using scan averaging
reduces the temporal resolution but maintains the maximum wavelength
resolution.)

```{r, eval=n>0L}
set_scans_to_avg(jwrapper = w, n.scans = 4L, sr.index = 0L, ch.index = 0L)
get_scans_to_avg(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

The correction for detector non-linearity can be applied in the spectrometer, 
but this can also be disabled. (According to Ocean Optics' literature the
in-spectrometer correction is applied based on the mean counts across the
whole array, which seems an odd thing to do. This needs checking as it
could distort the shape of the spectrum compared to applying the correction
pixel by pixel.)

```{r, eval=n>0L}
set_correct_for_detector_nonlinearity(jwrapper = w, correct = 0L, sr.index = 0L, ch.index = 0L)
get_correct_for_detector_nonlineary(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

This is another correction that can be applied in the spectrometer, based
on the reading of a small set of array pixels which are always darkened.

```{r, eval=n>0L}
set_correct_for_electrical_dark(jwrapper = w, enable = 0L, sr.index = 0L)
get_correct_for_electrical_dark(jwrapper = w, sr.index = 0L)
```

The time out is needed only with external triggering, and is the maximum
time to wait for a trigger before unblocking the communication with the
computer (the spectrometer becoming responsive to new commands).

```{r, eval=n>0L}
set_timeout(jwrapper = w, time.millisec = 1000L, sr.index = 0L)
```

Finally we acquire a spectrum! Subsequently we query whether it is valid,
and whether it is saturated (the sensor is clipping the peaks).

```{r, eval=n>0L}
raw.spectrum <- get_spectrum(jwrapper = w, sr.index = 0L, ch.index = 0L)
head(raw.spectrum)
summary(raw.spectrum)
is_spectrum_valid(jwrapper = w, sr.index = 0L, ch.index = 0L)
is_saturated(jwrapper = w, sr.index = 0L, ch.index = 0L)
```

The last step is always to close the connection to the spectrometers.

```{r, eval=n>0L}
close_all_spectrometers(w)
```

